---
title: "Correlation Analyses for COVID-19 Indicators"
author: "Delphi Group"
date: today
format:
  html:
    code-fold: true
---

```{r setup, include = FALSE}
# Core libraries
library(epidatr)
library(epiprocess)
library(covidcast) # Used for specialized plotting
library(dplyr)
library(ggplot2)
library(purrr)
library(tidyr)
library(lubridate)
library(scales)

# Set a modern, clean minimal theme for all plots
theme_set(
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "grey30"),
    panel.grid.minor = element_blank(),
    strip.background = element_blank()
  )
)

# Global knitr options
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  fig.width = 10, 
  fig.height = 7,
  dpi = 300,
  out.width = "100%"
)
```

## County analysis

First, we'll look at counties. 

### Getting data 

We fetch various signals from our API, from April 15 through to the current day.

```{r}

# Fetch the following sources and signals from the API 
sources <- c("doctor-visits", "fb-survey", "fb-survey", "hospital-admissions", 
            "indicator-combination")
signals <- c("smoothed_adj_cli", "smoothed_cli", "smoothed_hh_cmnty_cli", 
            "smoothed_adj_covid19_from_claims", "nmf_day_doc_fbc_fbs_ght")
names <- c("Doctor visits", "Facebook CLI", "Facebook CLI-in-community", 
          "Hospitalizations", "Combo indicator")

start_day = "2020-04-15"
end_day = as.character(Sys.Date())

df_signals = vector("list", length(signals))
for (i in 1:length(signals)) {
  # We use pub_covidcast and fetch, then convert to epi_df
  df_signals[[i]] <- pub_covidcast(sources[i], signals[i], "county", "day", 
                                  time_values = epirange(start_day, end_day)) %>%
                      as_epi_df()
}

# Fetch USAFacts confirmed case incidence proportion (smoothed with 7-day 
# trailing average)
df_cases <- pub_covidcast("usa-facts", "confirmed_7dav_incidence_prop",
                            "county", "day",
                            time_values = epirange(start_day, end_day)) %>%
             as_epi_df()
```

### Correlations sliced by time

Here we look at Spearman (rank) correlations between our signals and COVID-19
case incidence rates, sliced by time. That is, for each day, we compute the 
correlation between each signal and COVID-19 case incidence rates, over all  
counties (with at least 500 cumulative cases). 

```{r}
# Consider only counties with at least 500 cumulative cases
case_num <- 500
geo_values <- pub_covidcast("usa-facts", "confirmed_cumulative_num", 
                           "county", "day",
                           time_values = epirange(max(df_cases$time_value), max(df_cases$time_value))) %>%
               as_epi_df() %>%
               filter(value >= case_num) %>% pull(geo_value)

df_cor <- vector("list", length(signals))
for (i in 1:length(signals)) {
  # Join signal and cases
  combined <- full_join(df_signals[[i]] %>% filter(geo_value %in% geo_values), 
                       df_cases %>% filter(geo_value %in% geo_values), 
                       by = c("geo_value", "time_value"))
  
  # Compute correlation
  df_cor[[i]] <- epi_cor(combined, value.x, value.y, cor_by = time_value, method = "spearman") %>%
                  rename(value = cor)
  df_cor[[i]]$signal = names[i]
}
df <- do.call(rbind, df_cor)

ggplot(df, aes(x = time_value, y = value)) +
  geom_line(aes(color = signal), size = 1) +
  scale_x_date(limits = range(df$time_value, na.rm = TRUE), 
               date_breaks = "3 months", date_labels = "%b %y") +
  scale_y_continuous(labels = scales::comma) +
  guides(color = guide_legend(nrow = 2)) +
  labs(title = "Correlation between signals and case rates",
       subtitle = sprintf("Over all counties with at least %i cumulative cases",
                          case_num), x = "Date", y = "Correlation") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

### Correlations sliced by county

Now we look at Spearman (rank) correlations between our signals and COVID-19
case incidence rates, sliced by county. That is, for each county (with at least
500 cumulative cases), we compute the correlation between each signal and 
COVID-19 case incidence rates, over all time.

```{r}
df_cor <- vector("list", length(signals))
for (i in 1:length(signals)) {
  # Join signal and cases
  combined <- full_join(df_signals[[i]] %>% filter(geo_value %in% geo_values), 
                       df_cases %>% filter(geo_value %in% geo_values), 
                       by = c("geo_value", "time_value"))
  
  # Compute correlation
  df_cor[[i]] <- epi_cor(combined, value.x, value.y, cor_by = geo_value, method = "spearman") %>%
                  rename(value = cor)
  df_cor[[i]]$signal <- names[i]
}
df <- do.call(rbind, df_cor)

ggplot(df, aes(value)) +
  geom_density(aes(color = signal, fill = signal), alpha = 0.4) +
  scale_x_continuous(limits = c(-1, 1)) +
  guides(color = guide_legend(nrow = 2)) +
  labs(title = "Correlation between signals and case rates",
       subtitle = sprintf("Over all counties with at least %i cumulative cases",
                          case_num), x = "Correlation", y = "Density") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

We can also look at choropleth maps to get a geographic sense of the correlation
distribution for each signal.

```{r}
# Set some fields, then plot choropleth maps using covidcast functionality
for (i in 1:length(signals)) {
  # We need to add metadata so covidcast::plot knows how to handle this.
  # It expects a specific structure including data_source and signal.
  df_cor[[i]] = df_cor[[i]] %>% 
    mutate(
      time_value = as.Date(start_day), 
      issue = as.Date(start_day),
      data_source = "correlations",
      signal = signals[i]
    )
  
  attributes(df_cor[[i]])$metadata$geo_type = "county"
  class(df_cor[[i]]) = c("covidcast_signal", "data.frame")
  
  print(plot(df_cor[[i]], range = c(-1, 1), choro_col = cm.colors(10),
             title = sprintf("Correlations for %s", names[i])))
}
```

## Metro area analysis

Next, we'll look at metro areas. 

### Getting data 

We fetch various signals from our API, from April 15 through to the current day.

```{r}
# Fetch the following sources and signals from the API 
sources <- c("doctor-visits", "fb-survey", "fb-survey", "ght",
            "hospital-admissions", "indicator-combination")
signals <- c("smoothed_adj_cli", "smoothed_cli", "smoothed_hh_cmnty_cli", 
            "smoothed_search", "smoothed_adj_covid19_from_claims", 
            "nmf_day_doc_fbc_fbs_ght")
names <- c("Doctor visits", "Facebook CLI", "Facebook CLI-in-community", 
          "Google trends", "Hospitalizations", "Combo indicator")

start_day <- "2020-04-15"
end_day <- as.character(Sys.Date())

df_signals <- vector("list", length(signals))
for (i in 1:length(signals)) {
  df_signals[[i]] <- pub_covidcast(sources[i], signals[i], "msa", "day", 
                                  time_values = epirange(start_day, end_day)) %>%
                      as_epi_df()
}

# Fetch USAFacts confirmed case incidence proportion (smoothed with 7-day
# trailing average)
df_cases <- pub_covidcast("usa-facts", "confirmed_7dav_incidence_prop",
                            "msa", "day",
                            time_values = epirange(start_day, end_day)) %>%
             as_epi_df()
```

### Correlations sliced by time

Here we look at Spearman (rank) correlations between our signals and COVID-19
case incidence rates, sliced by time. That is, for each day, we compute the 
correlation between each signal and COVID-19 case incidence rates, over all 
metro areas (with at least 500 cumulative cases). 

```{r}
# Consider only metro areas with at least 500 cumulative cases
case_num <- 500
geo_values <- pub_covidcast("usa-facts", "confirmed_cumulative_num",
                               "msa", "day",
                               time_values = epirange(max(df_cases$time_value), max(df_cases$time_value))) %>%
  as_epi_df() %>%
  filter(value >= case_num) %>% pull(geo_value)

df_cor <- vector("list", length(signals))
for (i in 1:length(signals)) {
  # Join signal and cases
  combined <- full_join(df_signals[[i]] %>% filter(geo_value %in% geo_values), 
                       df_cases %>% filter(geo_value %in% geo_values), 
                       by = c("geo_value", "time_value"))
  
  # Compute correlation
  df_cor[[i]] <- epi_cor(combined, value.x, value.y, cor_by = time_value, method = "spearman") %>%
                  rename(value = cor)
  df_cor[[i]]$signal <- names[i]
}
df <- do.call(rbind, df_cor)

ggplot(df, aes(x = time_value, y = value)) +
  geom_line(aes(color = signal), size = 1) +
  scale_x_date(limits = range(df$time_value, na.rm = TRUE), 
               date_breaks = "3 months", date_labels = "%b %y") +
  scale_y_continuous(labels = scales::comma) +
  guides(color = guide_legend(nrow = 2)) +
  labs(title = "Correlation between signals and case rates",
       subtitle = sprintf("Over metro areas with at least %i cumulative cases",
                          case_num), x = "Date", y = "Correlation") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

### Correlations sliced by metro area

Now we look at Spearman (rank) correlations between our signals and COVID-19
case incidence rates, sliced by metro area. That is, for each metro area (with 
at least 500 cumulative cases), we compute the correlation between each signal 
and COVID-19 case incidence rates, over all time.

```{r}
df_cor <- vector("list", length(signals))
for (i in 1:length(signals)) {
  # Join signal and cases
  combined <- full_join(df_signals[[i]] %>% filter(geo_value %in% geo_values), 
                       df_cases %>% filter(geo_value %in% geo_values), 
                       by = c("geo_value", "time_value"))
  
  # Compute correlation
  df_cor[[i]] <- epi_cor(combined, value.x, value.y, cor_by = geo_value, method = "spearman") %>%
                  rename(value = cor)
  df_cor[[i]]$signal <- names[i]
}
df <- do.call(rbind, df_cor)

ggplot(df, aes(value)) +
  geom_density(aes(color = signal, fill = signal), alpha = 0.4) +
  scale_x_continuous(limits = c(-1, 1)) +
  guides(color = guide_legend(nrow = 2)) +
  labs(title = "Correlation between signals and case rates",
       subtitle = sprintf("Over metro areas with at least %i cumulative cases",
                          case_num), x = "Correlation", y = "Density") +
  theme(legend.position = "bottom", legend.title = element_blank())
```
