---
title: "Quidel dashboard"
author: "Delphi Lab"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
---

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

### Response Volume

This is a check on overall volume of raw results. There may be weekly effects but we would not expect to see any big systematic changes.

```{r, fig.width = 7, fig.height = 5}
library(covidcast)
library(dplyr)
library(ggplot2)

date_scale <-
  scale_x_date(
    date_breaks = "1 month",
    date_minor_breaks = "1 week",
    date_labels = "%b %Y"
  )

twenty_eight_days_ago = Sys.Date() - 28

# Sampling volume
df_quidel_states = covidcast_signal("quidel",
                                    "covid_ag_raw_pct_positive",
                                    start_day = twenty_eight_days_ago,
                                    geo_type = "state")
n_per_day = df_quidel_states %>%
  group_by(time_value) %>%
  summarize(n = sum(sample_size))

ggplot(n_per_day, aes(x = time_value, y = n)) +
  geom_line() + geom_point() + theme_bw() +
  labs(
    x = "Date",
    y = "Number of Responses",
    title = sprintf("Total responses: %i, mean per day: %i",
                    round(sum(n_per_day$n)), round(mean(n_per_day$n)))
  ) +
  date_scale
```

### Coverage

This measures how much state coverage we have in the samples (i.e., how many unique states are present each day), and how it has recently changed over time.

```{r, fig.width = 7, fig.height = 5}
# Sampling coverage
df_quidel_states = covidcast_signal("quidel",
                                    "covid_ag_raw_pct_positive",
                                    start_day = twenty_eight_days_ago,
                                    geo_type = "state")
states_per_day = df_quidel_states %>%
  group_by(time_value) %>%
  summarize(n = n())

ggplot(states_per_day, aes(x = time_value, y = n)) +
  geom_line() + geom_point() + theme_bw() +
  labs(
    x = "Date",
    y = "Number of States",
    title = sprintf(
      "Unique States: %i, mean per day: %i",
      length(unique(df_quidel_states$geo_value)),
      round(mean(states_per_day$n))
    )
  ) +
  date_scale
```

## State Coverage Map

This visualizes the state coverage -- how frequently does each state show up in the data over the last 28 days?

```{r, fig.width = 10, fig.height = 8}

source("dashboard_functions.R")
plot_28_day_frequency_state(df_quidel_states)
```