---
title: "Correlation Analyses for COVID-19 Indicators"
author: "Delphi Group"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
---

```{r, include = FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 7)
```

## County analysis

TODO, write something

### Getting data 

TODO, write something

```{r, message = FALSE}
library(covidcast)
library(dplyr)
library(ggplot2)

# Fetch the following sources and signals from the API 
sources = c("doctor-visits", "fb-survey", "fb-survey", "hospital-admissions", 
            "indicator-combination")
signals = c("smoothed_adj_cli", "smoothed_cli", "smoothed_hh_cmnty_cli", 
            "smoothed_adj_covid19", "nmf_day_doc_fbc_fbs_ght")
names = c("Doctor visits", "Facebook CLI", "Facebook CLI-in-community", 
          "Hospitalizations", "Combo indicator")

start_day = "2020-04-15"
end_day = NULL

df_signals = vector("list", length(signals))
for (i in 1:length(signals)) {
  df_signals[[i]] = covidcast_signal(sources[i], signals[i], start_day, end_day)
}

# Fetch USAFacts confirmed case incidence proportion (smoothed with 7-day 
# trailing average)
df_cases = covidcast_signal("usa-facts", "confirmed_7dav_incidence_prop",
                            start_day, end_day)
```

### Correlations sliced by time

TODO, write something

```{r, message = FALSE, warning = FALSE}
# Consider only counties with at least 500 cumulative cases
case_num = 500
geo_values = covidcast_signal("usa-facts", "confirmed_cumulative_num",
                              max(df_cases$time_value), 
                              max(df_cases$time_value)) %>%
  filter(value >= case_num) %>% pull(geo_value)

df_cor = vector("list", length(signals))
for (i in 1:length(signals)) {
  df_cor[[i]] = covidcast_cor(df_signals[[i]] %>% 
                                filter(geo_value %in% geo_values), 
                              df_cases %>% 
                                filter(geo_value %in% geo_values), 
                              by = "time_value", method = "spearman")
  df_cor[[i]]$signal = names[i]
}
df = do.call(rbind, df_cor)

ggplot(df, aes(x = time_value, y = value)) +
  geom_line(aes(color = signal)) +
  guides(color = guide_legend(nrow = 2)) +
  labs(title = "Correlation between signals and case rates",
       subtitle = sprintf("Over all counties with at least %i cumulative cases",
                          case_num), x = "Date", y = "Correlation") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

### Correlations sliced by county

TODO, write something

```{r, message = FALSE, warning = FALSE}
df_cor = vector("list", length(signals))
for (i in 1:length(signals)) {
  df_cor[[i]] = covidcast_cor(df_signals[[i]] %>% 
                                filter(geo_value %in% geo_values), 
                              df_cases %>% 
                                filter(geo_value %in% geo_values), 
                              by = "geo_value", method = "spearman")
  df_cor[[i]]$signal = names[i]
}
df = do.call(rbind, df_cor)

ggplot(df, aes(value)) +
  geom_density(aes(color = signal, fill = signal), alpha = 0.4) +
  guides(color = guide_legend(nrow = 2)) +
  labs(title = "Correlation between signals and case rates",
       subtitle = sprintf("Over all counties with at least %i cumulative cases",
                          case_num), x = "Date", y = "Correlation") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

TODO, write something

```{r}
# Set some fields, then plot choropleth maps using covidcast functionality
for (i in 1:length(signals)) {
  df_cor[[i]]$time_value = start_day
  df_cor[[i]]$issue = start_day
  attributes(df_cor[[i]])$geo_type = "county"
  class(df_cor[[i]]) = c("covidcast_signal", "data.frame")
  
  print(plot(df_cor[[i]], range = c(-1, 1), choro_col = cm.colors(10),
             title = sprintf("Correlations for %s", names[i])))
}
```

## Metro area analysis

TODO, write something

### Getting data 

TODO, write something

```{r, message = FALSE}
# Fetch the following sources and signals from the API 
sources = c("doctor-visits", "fb-survey", "fb-survey", "ght",
            "hospital-admissions", "indicator-combination")
signals = c("smoothed_adj_cli", "smoothed_cli", "smoothed_hh_cmnty_cli", 
            "smoothed_search", "smoothed_adj_covid19", 
            "nmf_day_doc_fbc_fbs_ght")
names = c("Doctor visits", "Facebook CLI", "Facebook CLI-in-community", 
          "Google trends", "Hospitalizations", "Combo indicator")

start_day = "2020-04-15"
end_day = NULL

df_signals = vector("list", length(signals))
for (i in 1:length(signals)) {
  df_signals[[i]] = covidcast_signal(sources[i], signals[i], start_day, end_day,
                                     geo_type = "msa")
}

# Fetch USAFacts confirmed case incidence proportion (smoothed with 7-day
# trailing average)
df_cases = covidcast_signal("usa-facts", "confirmed_7dav_incidence_prop",
                            start_day, end_day, geo_type = "msa")
```

### Correlations sliced by time

TODO, write something

```{r, message = FALSE, warning = FALSE}
# Consider only metro areas with at least 500 cumulative cases
case_num = 500
geo_values = covidcast_signal("usa-facts", "confirmed_cumulative_num",
                              max(df_cases$time_value), 
                              max(df_cases$time_value), 
                              geo_type = "msa") %>%
  filter(value >= case_num) %>% pull(geo_value)

df_cor = vector("list", length(signals))
for (i in 1:length(signals)) {
  df_cor[[i]] = covidcast_cor(df_signals[[i]] %>% 
                                filter(geo_value %in% geo_values), 
                              df_cases %>% 
                                filter(geo_value %in% geo_values), 
                              by = "time_value", method = "spearman")
  df_cor[[i]]$signal = names[i]
}
df = do.call(rbind, df_cor)

ggplot(df, aes(x = time_value, y = value)) +
  geom_line(aes(color = signal)) +
  guides(color = guide_legend(nrow = 2)) +
  labs(title = "Correlation between signals and case rates",
       subtitle = sprintf("Over metro areas with at least %i cumulative cases",
                          case_num), x = "Date", y = "Correlation") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

### Correlations sliced by metro area

TODO, write something

```{r, message = FALSE, warning = FALSE}
df_cor = vector("list", length(signals))
for (i in 1:length(signals)) {
  df_cor[[i]] = covidcast_cor(df_signals[[i]] %>% 
                                filter(geo_value %in% geo_values), 
                              df_cases %>% 
                                filter(geo_value %in% geo_values), 
                              by = "geo_value", method = "spearman")
  df_cor[[i]]$signal = names[i]
}
df = do.call(rbind, df_cor)

ggplot(df, aes(value)) +
  geom_density(aes(color = signal, fill = signal), alpha = 0.4) +
  guides(color = guide_legend(nrow = 2)) +
  labs(title = "Correlation between signals and case rates",
       subtitle = sprintf("Over metro areas with at least %i cumulative cases",
                          case_num), x = "Date", y = "Correlation") +
  theme(legend.position = "bottom", legend.title = element_blank())
```