---
title: 1. Compute percentage change over time
description: Compute percentage change of signal values over time.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Compute percentage change over time}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

A basic way of assessing growth in a signal is to look at its percentage change
over two neighboring time windows. We investigate this in the current vignette,
using the `pct_change()` function in the `modeltools` package. As in the getting
started guide, we focus on state-level COVID-19 case rates from the USAFacts 
data source, smoothed via 7-day trailing averages, from June to mid November. 

```{r}
library(covidcast)

start_day <- "2020-06-01"
end_day <- "2020-11-15"
geo_values <- c("ca", "fl", "ny", "tx")

case_rates <- suppressMessages(
  covidcast_signal(data_source = "usa-facts", 
                   signal = "confirmed_7dav_incidence_prop",
                   start_day = start_day, end_day = end_day, 
                   geo_type = "state", geo_values = geo_values))

summary(case_rates)
```

## Percentage change

The `pct_change()` function operates on a `covidcast_signal` data frame, and 
takes an argument `n`, indicating the size of the local window (in days) to use. 
For example, if `n = 10`, then to compute the percentage change on November 10, 
we use 100 * (B - A) / A, where A is the sum of the values between November 6 
and November 10, and A is the sum of the values between November 1 and November 
5. The default is `n = 14`, giving the percentage change between back-to-back
weeks.

```{r, message = FALSE, warning = FALSE}
library(modeltools)
library(dplyr)

case_rates <- pct_change(case_rates, n = 14)

case_rates %>% 
  arrange(geo_value) %>% 
  select(geo_value, time_value, value, pct_change) %>%
  print(n = 21)
```

We can see that a column `pct_change` column has been appended to the output 
data frame, which contains the percentage change values estimates. Next we plot 
these values alongside the signal itself. 

```{r, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 4}
library(ggplot2)
library(gridExtra)

state = "fl"

p1 <- ggplot(case_rates %>% filter(geo_value == state),
             aes(x = time_value, y = value)) +
  geom_line() + 
  labs(x = "Date", y = "Cases per 100,00 people") 

p2 <- ggplot(case_rates %>% filter(geo_value == state),
             aes(x = time_value, y = pct_change)) +
  geom_line() + 
  labs(x = "Date", y = "Weekly percentage change") 

grid.arrange(p1, p2, nrow = 1)
```

## Smoothing via 7-day averaging

Computing the percentage change between back-to-back days, using `n = 2`, will
generally return a pretty volatile sequence; below we show how to smooth this 
sequence by post-applying a 7-day trailing average, with a suitable application 
of `slide_by_geo()`. 

```{r, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 4}
case_rates <- case_rates %>%
  pct_change(n = 2, col_name = "pct_change_daily") %>%
  slide_by_geo(~ Mean(.x$pct_change_daily), n = 7, 
               col_name = "pct_change_daily_7dav")

case_rates %>% 
  arrange(geo_value) %>% 
  select(geo_value, time_value, value, pct_change_daily, 
         pct_change_daily_7dav) %>%
  head(n = 7)

p1 <- ggplot(case_rates %>% filter(geo_value == state),
             aes(x = time_value, y = value)) +
  geom_line() + 
  labs(x = "Date", y = "Cases per 100,00 people") 

p2 <- ggplot(case_rates %>% filter(geo_value == state),
             aes(x = time_value)) +
  geom_line(aes(y = pct_change_daily), col = "red") + 
  geom_line(aes(y = pct_change_daily_7dav), col = "blue") + 
  labs(x = "Date", y = "Daily percentage change") 

grid.arrange(p1, p2, nrow = 1)
```