---
title: Getting started with covidcastR
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with covidcastR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package exposes only one function: the `?covidcast_signal` function allows
users to retrieve a data frame (tibble) of values from the [COVIDcast
API](https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html). Using this
function, you can fetch any data you may be interested in analyzing, then use
other R packages to graph results, make maps, fit models, or join the data with
data from other sources.

## Installing

This package is not on CRAN yet, so it can be installed using the
[`devtools`](https://cran.r-project.org/package=devtools) package:

```{r, eval=FALSE}
devtools::install_github("cmu-delphi/covidcast", ref = "main",
                         subdir = "clients/covidcastR")
```

## Example use

For full usage information, see the function documentation at
`help(covidcast_signal)`.

To obtain smoothed estimates of COVID-like illness from our Facebook survey for
every county in the United States between 2020-05-01 and 2020-05-07:

```{r}
library(covidcastR)

suppressMessages(
    covidcast_signal("fb-survey", "smoothed_cli", start_day = "20200501",
                     end_day = "20200507")
)
```

Notice the use of `?suppressMessages` to hide progress reporting from the
function as it downloads the data; if you download particularly large amounts of
data, you may prefer to allow the progress reporting so you know how long to
wait.

Note that `geo_value` is the FIPS code for each reported county; counties with
insufficient sample size are not reported. The [API
documentation](https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html)
specifies how geographies are identified.

To request estimates for states instead of counties:

```{r}
suppressMessages(
    covidcast_signal("fb-survey", "smoothed_cli", start_day = "20200501",
                     end_day = "20200507", geo_type = "state")
)
```
