---
title: Getting started with covidcastR
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with covidcastR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package provides access to data frames of values from the [COVIDcast
API](https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html). Using the
`covidcast_signal()` function, you can fetch any data you may be interested in
analyzing, then use other R packages to graph results, make maps, fit models, or
join the data with data from other sources.

## Installing

This package is not on CRAN yet, so it can be installed using the
[`devtools`](https://cran.r-project.org/package=devtools) package:

```{r, eval=FALSE}
devtools::install_github("cmu-delphi/covidcast", ref = "main",
                         subdir = "R-packages/covidcast")
```

## Basic examples

For full usage information, see the [function
documentation](../reference/index.html).

To obtain smoothed estimates of COVID-like illness from our Facebook survey for
every county in the United States between 2020-05-01 and 2020-05-07, we can use
`covidcast_signal()`:

```{r}
library(covidcastR)

cli <- suppressMessages(
    covidcast_signal("fb-survey", "smoothed_cli", start_day = "20200501",
                     end_day = "20200507", geo_type = "county")
)
head(cli)
```

Each row represents one observation in one county on one day. The county FIPS
code is given in the `geo_value` column, the date in the `time_value` column.
Here `value` is the requested signal---in this case, the smoothed estimate of
the percentage of people with COVID-like illness, based on the symptom surveys.
`stderr` is its standard error. See the `covidcast_signal()` documentation for
details on the returned data frame.

Notice the use of `suppressMessages()` to hide progress reporting from the
function as it downloads the data; if you download particularly large amounts of
data, you may prefer to allow the progress reporting so you know how long to
wait.

Note that `geo_value` is the FIPS code for each reported county; counties with
insufficient sample size are not reported. The [API
documentation](https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html)
specifies how geographies are identified.

To request estimates for states instead of counties:

```{r}
cli <- suppressMessages(
    covidcast_signal("fb-survey", "smoothed_cli", start_day = "20200501",
                     end_day = "20200507", geo_type = "state")
)
head(cli)
```

One can also select a specific geographic region by its ID. For example, this is
the FIPS code for Allegheny County, Pennsylvania:

```{r}
cli <- suppressMessages(
    covidcast_signal("fb-survey", "smoothed_cli", start_day = "20200501",
                     end_day = "20200507", geo_type = "county",
                     geo_value = "42003")
)
head(cli)
```

If we are interested in exploring the available signals and their metadata, we
can use `covidcast_meta()` to fetch a data frame of the available signals:

```{r}
meta <- covidcast_meta()
head(meta)
```

This can be filtered and sliced as desired to obtain information about any
signal of interest.


## Mapping the data

The [`usmap` package](https://cran.r-project.org/package=usmap) can map counties
and states in the United States. In combination with `covidcastR`, one can
easily make maps of values on a specific date.

For example, on June 1, 2020, this is a map of the fraction of doctor's visits
in each state (as reported by our healthcare partners) that were due to
COVID-like symptoms:

```{r us_dv_map}
library(usmap)
library(ggplot2)

dv <- suppressMessages(
    covidcast_signal("doctor-visits", "smoothed_adj_cli",
                     start_day = "20200601", end_day = "20200601",
                     geo_type = "state")
)

## plot_usmap expects a "state" column
dv$state <- dv$geo_value

plot_usmap(data = dv, values = "value") +
    scale_fill_continuous(low = "white", high = "red", name = "% of visits")
```

(The warning messages appear to be an internal issue with the `usmap` package.)

It should also be possible to use the data with any other mapping packages you
prefer, as long as they can easily identify states, counties, MSAs, or the other
regions reported in the API according to the IDs reported in the API.

## Graphing the data

Using [`ggplot2`](https://ggplot2.tidyverse.org/) or your favorite plotting
package, we can also easily plot time series from specific regions.

For example, this plot shows the same value -- the fraction of doctor's visits
due to COVID-like symptoms -- for the states of New York and Pennsylvania from
March 1 through June 1, 2020.

```{r ny_time_series}
ny <- suppressMessages(
    covidcast_signal("doctor-visits", "smoothed_adj_cli",
                     start_day = "20200301", end_day = "20200601",
                     geo_type = "state", geo_values = c("pa", "ny"))
)

ggplot(ny, aes(x = time_value, y = value, color = geo_value)) +
    geom_line() +
    labs(x = "Date", y = "% of visits due to CLI",
         title = "Doctor's visits", color = "State")
```

This graph makes the peaks of the pandemic apparent; by loading other data
sources from the API, such as cases and deaths, we could begin to conduct
analyses comparing the time series against each other.
