---
title: Getting started with covidcast
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with covidcast}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package provides access to data frames of values from the [COVIDcast
API](https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html). Using the
`covidcast_signal()` function, you can fetch any data you may be interested in
analyzing, then use `plot.covidcast_signal()` to make plots and maps. Since the
data is provided as a simple data frame, you can also wrangle it into whatever
form you need to conduct your desired analyses using other packages and
functions.

## Installing

This package is not on CRAN yet, so it can be installed using the
[`devtools`](https://cran.r-project.org/package=devtools) package:

```{r, eval=FALSE}
devtools::install_github("cmu-delphi/covidcast", ref = "main",
                         subdir = "R-packages/covidcast")
```

## Basic examples

For full usage information, see the [function
documentation](../reference/index.html).

To obtain smoothed estimates of COVID-like illness from our Facebook survey for
every county in the United States between 2020-05-01 and 2020-05-07, we can use
`covidcast_signal()`:

```{r}
library(covidcast)

cli <- suppressMessages(
    covidcast_signal("fb-survey", "smoothed_cli", start_day = "2020-05-01",
                     end_day = "2020-05-07", geo_type = "county")
)
head(cli)
```

Each row represents one observation in one county on one day. The county FIPS
code is given in the `geo_value` column, the date in the `time_value` column.
Here `value` is the requested signal---in this case, the smoothed estimate of
the percentage of people with COVID-like illness, based on the symptom surveys.
`stderr` is its standard error. See the `covidcast_signal()` documentation for
details on the returned data frame.

Notice the use of `suppressMessages()` to hide progress reporting from the
function as it downloads the data; if you download particularly large amounts of
data, you may prefer to allow the progress reporting so you know how long to
wait.

Note that `geo_value` is the FIPS code for each reported county; counties with
insufficient sample size are not reported. The [API
documentation](https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html)
specifies how geographies are identified.

To request estimates for states instead of counties:

```{r}
cli <- suppressMessages(
    covidcast_signal("fb-survey", "smoothed_cli", start_day = "2020-05-01",
                     end_day = "2020-05-07", geo_type = "state")
)
head(cli)
```

One can also select a specific geographic region by its ID. For example, this is
the FIPS code for Allegheny County, Pennsylvania:

```{r}
cli <- suppressMessages(
    covidcast_signal("fb-survey", "smoothed_cli", start_day = "2020-05-01",
                     end_day = "2020-05-07", geo_type = "county",
                     geo_value = "42003")
)
head(cli)
```

This package provides convenient functions for plotting and mapping these
signals. For more details and examples, see `vignette("plotting-signals")`.

### Finding counties and metro areas

The COVIDcast API identifies counties by their 5-digit FIPS code and
metropolitan areas by their CBSA ID number. This means that to query a specific
county or metropolitan area, we must have some way to quickly find its
identifier.

This package includes several datasets intended to make the process easier. For
example, if we look at `?county_census`, we find that the package provides
census data on every county in the United States, including its FIPS code. To
find Allegheny County, I could search the data frame however is most convenient
to me:

```{r, message=FALSE}
library(dplyr)

county_census %>%
    filter(CTYNAME == "Allegheny County") %>%
    select(FIPS, CTYNAME, STNAME, POPESTIMATE2019)
```

Hence we see that 42003 is the code to provide to `geo_value` to select this
county. We also see we can obtain a 2019 population estimate, which could be
useful in your analyses to normalize or scale values by population.

Similarly, to find the Pittsburgh metropolitan area, we can use the
`?msa_census` dataset provided in the package:

```{r}
msa_census %>%
    filter(startsWith(NAME, "Pittsburgh")) %>%
    select(CBSA, NAME, LSAD, POPESTIMATE2019)
```

We can see that the Pittsburgh metropolitan area has CBSA ID 38300, and we can
also get its 2019 census-estimated population. We could pass this ID to
`covidcast_signal()` when using `geo_type = 'msa'`.

**Note:** The `msa_census` dataset includes types of area beyond metropolitan
statistical areas, including micropolitan statistical areas. The `LSAD` column
identifies the type of each area. The COVIDcast API only provides estimates for
metropolitan statistical areas, not for their divisions or for micropolitan
areas.

## Signal metadata

If we are interested in exploring the available signals and their metadata, we
can use `covidcast_meta()` to fetch a data frame of the available signals:

```{r}
meta <- covidcast_meta()
head(meta)
```

This can be filtered and sliced as desired to obtain information about any
signal of interest. The `covidcast_meta()` documentation describes the columns
and their meanings.
