---
title: Plotting and mapping signals
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Plotting and mapping signals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Once you've fetched some COVIDcast signals using `covidcast_signal()`, the
returned `covidcast` objects can be plotted and mapped in various ways. The data
structure is designed to be tidy and easily wrangled using your favorite
packages, but the covidcast package also provides some tools for plotting and
mapping signals in an easy way.

For this vignette, we'll use our [combination
signal](https://cmu-delphi.github.io/delphi-epidata/api/covidcast-signals/indicator-combination.html)
as an example; the combination indicator is a statistical combination of several
data sources collected by Delphi, and for every county provides a measure of
factors related to COVID activity. We'll also use incident case counts. Fetching
the data is simple:

```{r, message=FALSE}
library(covidcast)

comb <- suppressMessages(
  covidcast_signal(data_source = "indicator-combination", 
                   signal = "nmf_day_doc_fbc_fbs_ght",
                   start_day = "2020-07-01", end_day = "2020-07-14")
)
summary(comb)

inum <- suppressMessages(
  covidcast_signal(data_source = "usa-facts",
                   signal = "confirmed_7dav_incidence_num",
                   start_day = "2020-07-01", end_day = "2020-07-14")
)
summary(inum)
```

## Choropleth maps

```{r, include = FALSE}
knitr::opts_chunk$set(fig.width = 10, fig.height = 8)
```

The default `plot` method for `covidcast_signal` objects, 
`plot.covidcast_signal()`, produces choropleth maps by using `ggplot2` and the
`usmap` package:

```{r}
plot(comb)
```

The color scheme is automatically chosen to be similar to that used on the
online [COVIDcast mapping tool](https://covidcast.cmu.edu). Also, by default,
this map shows the most recent day of data available in the data frame. One can
choose the day and also choose the color scales, transparency level for mega 
counties, and title:

```{r}
plot(comb, time_value = "2020-07-04", choro_col = cm.colors(10), alpha = 0.4,
     title = "Combination of COVID-19 indicators on 2020-07-04")
```

By providing `breaks` and `colors`, we can create custom color scales, for
example to have a log-spaced color scale for incident case counts:

```{r}
breaks <- c(0, 1, 2, 5, 10, 20, 50, 100, 200)
colors <- c("#D3D3D3", "#FFFFCC", "#FEDDA2", "#FDBB79", "#FD9950", "#EB7538", 
            "#C74E32", "#A3272C", "#800026")

# Note that length(breaks) == length(colors) by design. This works as follows: 
# we assign colors[i] iff the value satisfies breaks[i] <= value < breaks[i+1], 
# where we take breaks[0] = -Inf and breaks[N+1] = Inf, for N = length(breaks)

plot(inum, choro_col = colors, choro_params = list(breaks = breaks),
     title = "New COVID cases (7-day trailing average) on 2020-07-14")
```

Lastly, we show how we can use custom breaks to (visually) answer the question: 
which counties have cumulative case rates of at least 1/100?

```{r}
cprop <- suppressMessages(
  covidcast_signal(data_source = "usa-facts",
                   signal = "confirmed_cumulative_prop",
                   start_day = "2020-07-01", end_day = "2020-07-14")
)

breaks <- c(0, 1000)
colors <- c("#D3D3D3", "#FFC0CB")

plot(cprop, choro_col = colors, 
     choro_params = list(breaks = breaks, legend_width = 3), 
     title = "Cumulative COVID cases per 100k people on 2020-07-14")
```

## Bubble maps

As an alternative to choropleth maps, we can also quickly plot bubble maps. By
default, bubble maps have 8 bubble size bins evenly spaced over the range, where
zero always means zero bubble size. The legend shows all bins, interpreted as
each bubble size meaning *at least* the corresponding value.

```{r}
plot(inum, plot_type = "bubble")
```

As before, we can of course set customized breaks. As values to the left of the
first bin do not get drawn, this map is much sparser, and highlights areas with
larger case counts.

```{r}
plot(inum, plot_type = "bubble", 
     bubble_params = list(breaks = seq(20, 200, len = 6)))
```

```{r, include = FALSE}
knitr::opts_chunk$set(fig.width = 10, fig.height = 6)
```

As a final example, suppose we want to plot only counties in the state of Texas.
We'd like to compare counts per 100,000 against absolute counts, so we fetch the
proportion signal:

```{r}
iprop <- suppressMessages(
  covidcast_signal(data_source = "usa-facts",
                   signal = "confirmed_7dav_incidence_prop",
                   start_day = "2020-07-01", end_day = "2020-07-14")
)
```

Then we make two maps side-by-side with custom ranges:

```{r}
library(gridExtra)

breaks1 <- c(1, 10, 100, 1000)
breaks2 <- c(10, 50, 100, 500)

p1 <- plot(inum, plot_type = "bubble", 
           bubble_params = list(breaks = breaks1, max_size = 6),
           include = "TX", bubble_col = "red",
           title = paste("Incidence number on", max(inum$time_value)))
p2 <- plot(iprop, plot_type = "bubble", 
           bubble_params = list(breaks = breaks2, max_size = 6),
           include = "TX", bubble_col = "red",
           title = paste("Incidence rate on", max(iprop$time_value)))

grid.arrange(p1, p2, nrow = 1)
```

## Time series plots

```{r, include = FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 6)
```

Let's fetch the combination indicator and case counts, but for all states rather
than for all counties. This will make the time series plots more manageable.

```{r}
suppressMessages({
comb_st <- covidcast_signal(data_source = "indicator-combination",
                               signal = "nmf_day_doc_fbc_fbs_ght",
                               start_day = "2020-04-15", end_day = "2020-07-01",
                               geo_type = "state")
inum_st <- covidcast_signal(data_source = "usa-facts",
                               signal = "confirmed_7dav_incidence_num",
                               start_day = "2020-04-15", end_day = "2020-07-01",
                               geo_type = "state")
})
```

By default, time series plots show all available data, including all
geographies. A line for every state would be unmanageable, so let's select a few
states and plot all data for them:

```{r, message = FALSE}
library(dplyr)

states <- c("ca", "pa", "tx", "ny")
plot(comb_st %>% filter(geo_value %in% states), plot_type = "line") 
plot(inum_st %>% filter(geo_value %in% states), plot_type = "line")
```

Notice how in Texas, the combined indicator rose several weeks in advance of
confirmed cases, suggesting the signal could be predictive. Delphi is
investigating these signals for their usefulness in forecasting, as well as 
hotspot detection and will publish results when they are available.

## Manual plotting

Using `ggplot2` or your favorite plotting package, we can easily plot time
series manually, without using the `plot.covidcast_signal()` method. You can use
this to customize the appearance of your plots however you choose.

For example:

```{r, warning = FALSE}
library(ggplot2)

suppressMessages({
comb_md <- covidcast_signal(data_source = "indicator-combination",
                               signal = "nmf_day_doc_fbc_fbs_ght",
                               start_day = "2020-06-01", end_day = "2020-07-15",
                               geo_values = name_to_fips("Miami-Dade"))
inum_md <- covidcast_signal(data_source = "usa-facts",
                               signal = "confirmed_7dav_incidence_num",
                               start_day = "2020-06-01", end_day = "2020-07-15",
                               geo_values = name_to_fips("Miami-Dade"))
})

# Compute the ranges of the two signals
range1 <- inum_md %>% select("value") %>% range
range2 <- comb_md %>% select("value") %>% range

# Function to transform from one range to another
trans <- function(x, from_range, to_range) {
  (x - from_range[1]) / (from_range[2] - from_range[1]) *
    (to_range[2] - to_range[1]) + to_range[1]
}

# Convenience functions for our two signal ranges
trans12 <- function(x) trans(x, range1, range2)
trans21 <- function(x) trans(x, range2, range1)

# Transform the combined signal to the incidence range, then stack
# these rowwise into one data frame
df <- select(rbind(comb_md %>% mutate_at("value", trans21),
                   inum_md), c("time_value", "value"))
df$signal <- c(rep("Combined indicator", nrow(comb_md)),
               rep("New COVID-19 cases", nrow(inum_md)))   

# Finally, plot both signals
ggplot(df, aes(x = time_value, y = value)) +
  labs(x = "Date", title = "Miami-Dade County") +
  geom_line(aes(color = signal)) +
  scale_y_continuous(
    name = "New COVID-19 cases (7-day trailing average)",
    sec.axis = sec_axis(trans12, name = "Combination of COVID-19 indicators")
  ) +
  theme(legend.position = "bottom",
        legend.title = ggplot2::element_blank())
```

Again, we see that the combined indicator starts rising several days before the
new COVID-19 cases do, an exciting phenomenon that Delphi is studying now.